---
title: "cytofkit CD14negCD19neg PBMC Analysis"
author: "Hao Chen"
date: "6 June 2016"
output: 
  html_document:
    toc: true
    toc_depth: 3
---
 
## Load the package

```{r}
require(cytofkit)  ## version >= 1.4.6
require(ggplot2)
require(reshape2)
require(plotly)
setwd("/Users/chenhao/GitProject/cytofkit_data_codes")  ## change this to your path
```

## Parameter setting and Data loading

```{r, eval=FALSE}
dimReductionMethod <- "tsne"
visualizationMethods <- "tsne"
clusterMethods <- c("Rphenograph", "ClusterX", "DensVM")
markers <- as.character(read.table("parameter.txt", header = TRUE)[ ,1])

xdata <- cytof_exprsExtract(fcsFile = "./CD14negCD19negPBMC_dataset/130515_C2_stim_CD19-.fcs", 
                           transformMethod = "autoLgcl",
                           markers = markers)
```


## Dimensionality reduction

```{r, eval=FALSE}
ydata <- cytof_dimReduction(data = xdata, method = "tsne")
```


## Cluster anlsysis using method Rphenograph, ClusterX, DensVM  

```{r, eval=FALSE}
cluster_res <- lapply(clusterMethods, cytof_cluster, 
                      ydata = ydata,
                      xdata = xdata)
names(cluster_res) <- clusterMethods
```


## Rename the clusters according to positions on t-SNE map for comparation.

```{r, eval=FALSE}
tsne_cluster <- as.data.frame(ydata)
tsne_cluster$ClusterX <- as.integer(cluster_res[["ClusterX"]])
tsne_cluster$DensVM <- as.integer(cluster_res[["DensVM"]])
tsne_cluster$PhenoGraph <- as.integer(cluster_res[["Rphenograph"]])

ClusterX_center <- aggregate(cbind(tsne_1, tsne_2) ~ ClusterX, data = tsne_cluster, median)
ClusterX_center$ClusterX_label <- rank(-ClusterX_center$tsne_2)
tsne_cluster$ClusterX <- ClusterX_center$ClusterX_label[match(tsne_cluster$ClusterX, ClusterX_center$ClusterX)]

DensVM_center <- aggregate(cbind(tsne_1, tsne_2) ~ DensVM, data = tsne_cluster, median)
DensVM_center$DensVM_label <- rank(-DensVM_center$tsne_2)
tsne_cluster$DensVM <- DensVM_center$DensVM_label[match(tsne_cluster$DensVM, DensVM_center$DensVM)]

PhenoGraph_center <- aggregate(cbind(tsne_1, tsne_2) ~ PhenoGraph, data = tsne_cluster, median)
PhenoGraph_center$PhenoGraph_label <- rank(-PhenoGraph_center$tsne_2)
tsne_cluster$PhenoGraph <- PhenoGraph_center$PhenoGraph_label[match(tsne_cluster$PhenoGraph, PhenoGraph_center$PhenoGraph)]

data_tsne_cluster <- cbind(xdata, tsne_cluster)
```


```{r}
## Sice t-SNE will output different embedding map every time, to reproduce the resutls in 
## cytofkit paper, we loaded the saved t-SNE and  cluster results from our run.
data_tsne_cluster <- read.csv("./CD14negCD19negPBMC_dataset/130515_C2_stim_CD19-_exprs_tsne_cluster.csv", header = TRUE, row.names = 1, check.names = FALSE)
```


## Comparison of clustering methods for subset detection

### PhenoGraph cluster plot
```{r}
cytof_clusterPlot(data_tsne_cluster, xlab = "tsne_1", ylab = "tsne_2", cluster = "PhenoGraph", sampleLabel = FALSE)
```


### ClusterX cluster plot
```{r}
cytof_clusterPlot(data_tsne_cluster, xlab = "tsne_1", ylab = "tsne_2", cluster = "ClusterX", sampleLabel = FALSE)
```


### DensVM cluster plot
```{r}
cytof_clusterPlot(data_tsne_cluster, xlab = "tsne_1", ylab = "tsne_2", cluster = "DensVM", sampleLabel = FALSE)
```

### FlowSOM cluster plot
```{r}
xdata <- data_tsne_cluster[,1:36]
```

```{r, cache=TRUE}
FlowSOM_15 <- cytof_cluster(xdata = xdata, method = "FlowSOM", FlowSOM_k = 15)
FlowSOM_14 <- cytof_cluster(xdata = xdata, method = "FlowSOM", FlowSOM_k = 14)
FlowSOM_13 <- cytof_cluster(xdata = xdata, method = "FlowSOM", FlowSOM_k = 13)
FlowSOM_10 <- cytof_cluster(xdata = xdata, method = "FlowSOM", FlowSOM_k = 10)
FlowSOM_5 <- cytof_cluster(xdata = xdata, method = "FlowSOM", FlowSOM_k = 5)

data_tsne_cluster$FlowSOM15 <- FlowSOM_15
data_tsne_cluster$FlowSOM14 <- FlowSOM_14
data_tsne_cluster$FlowSOM13 <- FlowSOM_13
data_tsne_cluster$FlowSOM10 <- FlowSOM_10
data_tsne_cluster$FlowSOM5 <- FlowSOM_5

## FlowSOM cluster k = 15
cytof_clusterPlot(data_tsne_cluster, xlab = "tsne_1", ylab = "tsne_2", cluster = "FlowSOM15", sampleLabel = FALSE)

## FlowSOM cluster k = 14
cytof_clusterPlot(data_tsne_cluster, xlab = "tsne_1", ylab = "tsne_2", cluster = "FlowSOM14", sampleLabel = FALSE)

## FlowSOM cluster k = 13
cytof_clusterPlot(data_tsne_cluster, xlab = "tsne_1", ylab = "tsne_2", cluster = "FlowSOM13", sampleLabel = FALSE)

## FlowSOM cluster k = 10
cytof_clusterPlot(data_tsne_cluster, xlab = "tsne_1", ylab = "tsne_2", cluster = "FlowSOM10", sampleLabel = FALSE)

## FlowSOM cluster k = 5
cytof_clusterPlot(data_tsne_cluster, xlab = "tsne_1", ylab = "tsne_2", cluster = "FlowSOM5", sampleLabel = FALSE)
```



## Cluster heatmap plot

### PhenoGraph cluster median heatmap
```{r}
cluster_median <- aggregate(. ~ PhenoGraph, data = data_tsne_cluster[,c(1:36,39)], median)
rownames(cluster_median) <- cluster_median$PhenoGraph
cluster_median <- subset(cluster_median, select = -PhenoGraph)
m <- regexpr("\\<.*\\>", colnames(cluster_median), perl = TRUE)
colnames(cluster_median) <- gsub("<|>", "", regmatches(colnames(cluster_median), m))
cytof_heatmap(cluster_median, baseName = "Cluster Median", cex_row_label = 1, cex_col_label = 0.8)
```


### ClusterX cluster median heatmap
```{r}
cluster_median <- aggregate(. ~ ClusterX, data = data_tsne_cluster[,c(1:36,40)], median)
rownames(cluster_median) <- cluster_median$ClusterX
cluster_median <- subset(cluster_median, select = -ClusterX)
m <- regexpr("\\<.*\\>", colnames(cluster_median), perl = TRUE)
colnames(cluster_median) <- gsub("<|>", "", regmatches(colnames(cluster_median), m))
cytof_heatmap(cluster_median, baseName = "Cluster Median", cex_row_label = 1, cex_col_label = 0.8)
```


### DensVM cluster median heatmap
```{r}
cluster_median <- aggregate(. ~ DensVM, data = data_tsne_cluster[,c(1:36,41)], median)
rownames(cluster_median) <- cluster_median$DensVM
cluster_median <- subset(cluster_median, select = -DensVM)
m <- regexpr("\\<.*\\>", colnames(cluster_median), perl = TRUE)
colnames(cluster_median) <- gsub("<|>", "", regmatches(colnames(cluster_median), m))
cytof_heatmap(cluster_median, baseName = "Cluster Median", cex_row_label = 1, cex_col_label = 0.8)
```


## Assess ISOMAP, diffusion map and t-SNE for inferring inter-cluster relationship

```{r}
## Loading subsample data
subData1 <- read.csv("./CD14negCD19negPBMC_dataset/130515_C2_stim_CD19-_subsample1_tsne.csv", header = TRUE, row.names = 1, check.names = FALSE)
subData2 <- read.csv("./CD14negCD19negPBMC_dataset/130515_C2_stim_CD19-_subsample2_tsne.csv", header = TRUE, row.names = 1, check.names = FALSE)
subData3 <- read.csv("./CD14negCD19negPBMC_dataset/130515_C2_stim_CD19-_subsample3_tsne.csv", header = TRUE, row.names = 1, check.names = FALSE)
```

### t-SNE plot

```{r}
cytof_clusterPlot(subData1, xlab = "tsne_1", ylab = "tsne_2", 
                  cluster = "cluster", sampleLabel = FALSE, 
                  labelSize = 8, fixCoord = FALSE)

cytof_clusterPlot(subData2, xlab = "tsne_1", ylab = "tsne_2", 
                  cluster = "cluster", sampleLabel = FALSE, 
                  labelSize = 8, fixCoord = FALSE)

cytof_clusterPlot(subData3, xlab = "tsne_1", ylab = "tsne_2", 
                  cluster = "cluster", sampleLabel = FALSE, 
                  labelSize = 8, fixCoord = FALSE)
```

### ISOMAP plot

```{r, cache=TRUE}
## Loading subsample data
subData1 <- read.csv("./CD14negCD19negPBMC_dataset/130515_C2_stim_CD19-_subsample1_isomap.csv", header = TRUE, row.names = 1, check.names = FALSE)
subData2 <- read.csv("./CD14negCD19negPBMC_dataset/130515_C2_stim_CD19-_subsample2_isomap.csv", header = TRUE, row.names = 1, check.names = FALSE)
subData3 <- read.csv("./CD14negCD19negPBMC_dataset/130515_C2_stim_CD19-_subsample3_isomap.csv", header = TRUE, row.names = 1, check.names = FALSE)
```

```{r}
cytof_clusterPlot(subData1, xlab = "isomap_1", ylab = "isomap_2", 
                  cluster = "cluster", sampleLabel = FALSE, 
                  labelSize = 8, fixCoord = FALSE)

cytof_clusterPlot(subData2, xlab = "isomap_1", ylab = "isomap_2", 
                  cluster = "cluster", sampleLabel = FALSE, 
                  labelSize = 8, fixCoord = FALSE)

cytof_clusterPlot(subData3, xlab = "isomap_1", ylab = "isomap_2", 
                  cluster = "cluster", sampleLabel = FALSE, 
                  labelSize = 8, fixCoord = FALSE)
```


### Diffusion map plot
```{r, cache=TRUE}
## diffusion map for subsample 1
dfmap1 <- cytof_progression(data = subData1[ ,1:36], 
                            method = "diffusionmap",
                            out_dim = 4,
                            cluster = subData1[,37],
                            clusterSampleMethod = "all")

## diffusion map for subsample 2
dfmap2 <- cytof_progression(data = subData2[ ,1:36], 
                            method = "diffusionmap",
                            out_dim = 4,
                            cluster = subData2[,37],
                            clusterSampleMethod = "all")

## diffusion map for subsample 3
dfmap3 <- cytof_progression(data = subData3[ ,1:36], 
                            method = "diffusionmap",
                            out_dim = 4,
                            cluster = subData3[,37],
                            clusterSampleMethod = "all")
```



```{r}
dfmap1_progression_data <- as.data.frame(do.call(cbind, dfmap1))
cytof_clusterPlot(dfmap1_progression_data, xlab = "diffusionmap_1", ylab = "diffusionmap_2", 
                  cluster = "sampleCluster", sampleLabel = FALSE, 
                  labelSize = 8, fixCoord = FALSE)


dfmap2_progression_data <- as.data.frame(do.call(cbind, dfmap2))
cytof_clusterPlot(dfmap2_progression_data, xlab = "diffusionmap_1", ylab = "diffusionmap_2", 
                  cluster = "sampleCluster", sampleLabel = FALSE, 
                  labelSize = 8, fixCoord = FALSE)

dfmap3_progression_data <- as.data.frame(do.call(cbind, dfmap3))
cytof_clusterPlot(dfmap3_progression_data, xlab = "diffusionmap_1", ylab = "diffusionmap_2", 
                  cluster = "sampleCluster", sampleLabel = FALSE, 
                  labelSize = 8, fixCoord = FALSE)
```



## Progression analysis

```{r, eval=FALSE}
## load some functions
isomap_progression <- cytof_progression(data = data_tsne_cluster[ ,1:36], 
                                        cluster = data_tsne_cluster$ClusterX,
                                        method = "isomap",
                                        out_dim = 4,
                                        clusterSampleMethod = "ceil",
                                        clusterSampleSize = 500)
isomap_progression_data <- as.data.frame(do.call(cbind, isomap_progression))
write.csv(isomap_progression_data, "subset_downsample_500_isomap_progression.csv")

dfmap_progression <- cytof_progression(data = data_tsne_cluster[ ,1:36], 
                                        method = "diffusionmap",
                                        out_dim = 4,
                                        cluster = data_tsne_cluster$ClusterX,
                                        clusterSampleMethod = "ceil",
                                        clusterSampleSize = 500)

dfmap_progression_data <- as.data.frame(do.call(cbind, dfmap_progression))
write.csv(dfmap_progression_data, "subset_downsample_500_diffusionmap_progression.csv")
```

```{r, echo=FALSE}
isomap_progression_data <- read.csv("./CD14negCD19negPBMC_dataset/subset_downsample_500_isomap_progression.csv", header = TRUE, row.names = 1, check.names = FALSE)

dfmap_progression_data <- read.csv("./CD14negCD19negPBMC_dataset/subset_downsample_500_diffusionmap_progression.csv", header = TRUE, row.names = 1, check.names = FALSE)
```


### Progression plot

```{r}
#plot_ly(isomap_progression_data, x = isomap_1, y = isomap_2, z = isomap_3, group = sampleCluster, size = 0.5, type = "scatter3d", mode = "markers")

cytof_clusterPlot(isomap_progression_data, xlab = "isomap_1", ylab = "isomap_2", 
                  cluster = "sampleCluster", sampleLabel = FALSE, 
                  labelSize = 15, fixCoord = FALSE)

#plot_ly(dfmap_progression_data, x = diffusionmap_1, y = diffusionmap_2, z = diffusionmap_3, group = sampleCluster, type = "scatter3d", mode = "markers")

cytof_clusterPlot(dfmap_progression_data, xlab = "diffusionmap_1", ylab = "diffusionmap_2", 
                  cluster = "sampleCluster", sampleLabel = FALSE, 
                  labelSize = 15, fixCoord = FALSE)

```


```{r, echo=FALSE}
## color plot funciton
remove_outliers <- function(x, na.rm = TRUE, ...) {
    qnt <- quantile(x, probs=c(.25, .75), na.rm = na.rm, ...)
    H <- 1.5 * IQR(x, na.rm = na.rm)
    y <- x
    y[x < (qnt[1] - H)] <- qnt[1] - H
    y[x > (qnt[2] + H)] <- qnt[2] + H
    y
}

visuaPlot <- function(data, xlab, ylab, zlab, pointSize=1, 
                      addLabel=TRUE, labelSize=1, removeOutlier = TRUE){
    
    data <- as.data.frame(data)
    title <- zlab
    data <- data[,c(xlab, ylab, zlab)]
    if(removeOutlier)
        data[,zlab] <- remove_outliers(data[,zlab])
    zlab <- "Expression"
    colnames(data) <- c(xlab, ylab, zlab)
    gp <- ggplot(data, aes_string(x = xlab, y = ylab, colour = zlab)) + 
        geom_point(size = pointSize, alpha = 1) + theme_bw() +
        scale_colour_gradient2(low="blue", mid="white", high="red",     midpoint=median(data[[zlab]])) +
        theme(legend.position = "right") + xlab(xlab) + ylab(ylab) + ggtitle(title) +
        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
        theme(axis.text=element_text(size=8), axis.title=element_text(size=12,face="bold"))
    
    return(gp)
}
```


```{r}
visMarkers <- c("(Sm150)Di<GranzymeB>", "(Yb173)Di<Perforin>")

visuaPlot(isomap_progression_data, xlab = "isomap_1", ylab="isomap_2", 
          zlab = visMarkers[1], pointSize = 1.5)

visuaPlot(isomap_progression_data, xlab = "isomap_1", ylab="isomap_2", 
          zlab = visMarkers[2], pointSize = 1.5)

visuaPlot(dfmap_progression_data, xlab = "diffusionmap_1", ylab="diffusionmap_2", 
          zlab = visMarkers[1], pointSize = 1.5)

visuaPlot(dfmap_progression_data, xlab = "diffusionmap_1", ylab="diffusionmap_2", 
          zlab = visMarkers[2], pointSize = 1.5)



cytof_progressionPlot(data=isomap_progression_data, 
                      markers = visMarkers, 
                      clusters = c(11, 14, 12, 15, 13), 
                      orderCol = "isomap_2",
                      clusterCol = "sampleCluster",
                      clusterLabelSize = 10,
                      segmentSize = 1,
                      reverseOrder = TRUE,
                      addClusterLabel = TRUE)


cytof_progressionPlot(data=dfmap_progression_data, 
                      markers = visMarkers, 
                      clusters = c(11, 14, 12, 15, 13), 
                      orderCol = "diffusionmap_2",
                      clusterCol = "sampleCluster",
                      clusterLabelSize = 10,
                      segmentSize = 1,
                      reverseOrder = TRUE,
                      addClusterLabel = TRUE)

```


## Session Information

```{r}
sessionInfo()
```

